\documentclass[a4paper]{report}

\usepackage{array}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage[bookmarks]{hyperref}
\usepackage{pgf,tikz}
\usepackage{mathrsfs}
\usepackage{amsmath}
\usetikzlibrary{arrows}
\usepackage{listings}
\usepackage{hyperref}
%\usepackage{multirow}

\input{diverse/glsl-listings.tex}

\begin{document}
\title{	Bachelor Project: \\
	\textbf{Using Deep G-buffers for Interactive Screenspace Global Illumination}}
\author{Jonas H. Nielsen, s121751}
\date{\parbox{\linewidth}{\centering%
  \today\endgraf\bigskip
  Supervisors\endgraf
  Niels Jørgen Christensen \hspace*{3cm} Jeppe Revall Frisvad\endgraf\medskip
  DTU Compute \endgraf
  Technical University of Denmark}}
\maketitle
\newpage
\begin{abstract}
The purpose of this thesis is to implement and explain a number of screen-space global illumination methods. In order to achieve this goal I employ a deep G-buffer which provides our screen-space geometric data with an additional layer to allow for more scene consistency. I start out by going through the theory behind deferred rendering and its advantages, and use that to explain the advantages and use of a deep G-buffer implementation. I also explain the theory behind an omni-directional shadow map implementation based on a cube map texture produced in a single pass. I provide the theory for radiosity by taking a point of departure in the rendering equation and arriving at a quasi-Monte Carlo based screen-space solution. Additionally I explain the Alchemy SSAO algorithm and a filtering method to even out the results of both SSAO and radiosity.

I describe the implementation of all concepts mentioned under theory by using code samples before moving on to my results. Based on the quality and running times of the results I conclude that the deep G-buffer, for the purpose of the scenes used in this thesis, does not provide enough advantages to justify its extra cost. I consider the implementation of all concepts successful, but radiosity and the omni-directional shadow map need optimisations to be viable. The use of the depth buffer to reconstruct the scene is concluded to be a viable way to trade off a small amount of performance to potentially free up a lot of space from the G-buffer.
\end{abstract}
\newpage
\tableofcontents
\newpage

\input{1introduction/intro.tex}
%\input{2hypethesis/hyp.tex}
\input{3theory/theory.tex}
\input{4implementation/impl.tex}
\input{5results/results.tex}
%\input{6discussion/disc.tex}
\input{7conclusion/conc.tex}
\input{bibliography/bib.tex}

\end{document}